---
title: "Bayesian Dose Response Modeling"
output: html_document
---

The code block below provides the code required to install the BayesPharma 
package.

```{r}
# install.packages("remotes")
# remotes::install_github("maomlab/BayesPharma")
```

Then, we need to load the library. All the dependencies packages will be loaded
as well.
```{r}
library(dplyr)
library(BayesPharma)
```


Load the clean dataset below. 
```{r}
clean_dr_data <- readr::read_csv("C:/Users/marti/Documents/Bayesetta_Workshop/dose_response_modeling/clean_bayesetta_dr_data.csv")
dr_data <- clean_dr_data
dr_data
```

To do a dose response relationship curve analysis, we need to have the 
concentration to be in the logarithmic base 10 scale. 
The BayesPharma package has a function to calculate that and add a column 
automatically. Use the `calculate_log_dose()` function to add a *log_dose* to 
the data.frame. Because the dose column is labelled "dose" 
and the dose values are in molar concentration we can use the defaults:
*dose_col = dose* and *molar_concentration = 1*.
To find information about the inputs for the BayesPharma functions, go to the 
'Packages' tab and click on "BayesPharma".
```{r}
dr_data <- dr_data %>% calculate_log_dose()
dr_data
```

The BayesPharma package comes with limitations, so we have to change the column 
names to be compatible with BayesPharma functions. The response column name must 
be "response" and the log dose column name must be "log_dose". The 
`calculate_log_dose()` function automatically labels the new log dose column 
"log_dose". 

Use the `change_col_names()`function to change the "percent_response" column 
name. 

```{r}
dr_data <- dr_data %>% change_col_names(response_col_name = percent_response)
dr_data
```

Time to get into the Bayesian analysis!
First, we will run the analysis with top (max response) parameter prior set to 
a constant value of 100 because top is normalized to 100 and the default weakly 
informative priors for ic50, hill and bottom. 

The level of informativeness of the priors will affect how much influence the
priors have on the model.

Use the `dr_priors()` function and set *top = 100*. 
```{r}
data_priors <- dr_priors(top = 100)
data_priors
```

Then, we will set inits to values close to parameter estimates to help with 
model convergence. The inits are the starting values for the mcmc chains of each 
parameter. The default inits: ec50 = -9, hill = -1, top = 100, bottom = 0.

```{r}
data_inits <- dr_inits()
```

Before running the model, we will verify that the prior distributions
cover a plausible range of values for each parameter. To do this, we want to 
sample only from the prior distributions by adding  *sample_prior = "only"* as 
an argument to the `dr_model()` function.
For the *formula* argument, we will use the BayesPharma `dr_formula()` function.
We can use the default argument for the formula for evaluating the prior. 
Set *data* to the dr_data, *priors* to the priors we defined above and *inits* 
as the inits we defined above.

```{r}
data_sample_priors <- dr_model(
  data = dr_data,
  formula = dr_formula(), 
  priors = data_priors,
  inits = data_inits,
  sample_prior = "only")
data_sample_priors
```

Plot the prior distributions using the `density_distributions()` function. We
can use this to evaluate the prior distributions for each parameter. 
```{r}
density_distributions(data_sample_priors,
                      half_max_label = "ic50",
                      title_label = "Prior Density Distributions")
```

The prior distributions cover a broad plausible range of values. Now we are 
going to fit the model!
---

We want to estimate the values separately for each drug, therefore, we need to
set *predictors = 0 + drug* for the `dr_formula()`. 
```{r}
data_model <- dr_model(
  data = dr_data,
  formula = dr_formula(predictors = 0 + drug), 
  priors = data_priors,
  inits = data_inits)
```



```{r}
data_model
```


The model ran without any warning messages, Rhat for all parameters is 1.00 
(NA for top because it was set to a constant value of 100), and Bulk and Tail 
ESS are high. We can check the mixing and convergence of the mcmc chains using 
`traceplot()`. Set *predictors_col_name = 'drug'* and *half_max_label = 'ic50'*.
```{r}
data_model %>% traceplot(predictors_col_name = 'drug',
                         half_max_label = "ic50")
```

The mcmc chains appear to have mixed well throughout the model run. Therefore, 
we are in good shape to move forward with the analysis.

---
We can plot a posterior predictive check to check if the data generated from the 
posterior distribution fits the observed data. Use `plot_pp_check()` and set 
*model* to our fitted model, *plot_type = "dens_overlay_grouped"* and
*group = drug*. 

The observed data is the dark blue line and the simulated data are the 
light blue lines.

```{r}
plot_pp_check(data_model,
              plot_type = "dens_overlay_grouped",
              group = 'drug')

```


Let's plot the prior and posterior distributions of the parameters (prior is 
pink and posterior is teal). this can be useful if you are interested in 
comparing the initial prior with the posterior distribution produced by the 
model. 
Include the fitted model, set *predictors_col_name = "drug"* and 
*half_max_label = "ic50"*.

```{r}
prior_posterior_densities(data_model,
                          predictors_col_name = "drug",
                          half_max_label = "ic50")
```

Let's plot  the posterior distributions for each parameter 
with the confidence intervals and mean. This is a useful visual of the model 
results. the same arguments as the plot above can be used here. 

```{r}
posterior_densities(data_model, 
                     predictors_col_name = "drug", 
                     half_max_label = "ic50")
```

Finally we will plot the dose response curves! The plot is a sample of 50 
sigmoid dose-response curves from the posterior distribution (purple) and 
the median quantile intervals. Set *model* as our fitted model, 
*data* to the dr_data, *predictors_col_name = "drug"*, *facet_var = drug*.

```{r}
posterior_draws_plot(model = data_model, data = dr_data, 
                     predictors_col_name = "drug", facet_var = drug)

# model, data, predictors_col_name = "na",
# lower = -12, upper = -3, n = 50,
# facet_var, point_size = 0.75,
# jitter_height = 0, jitter_width = 0,
# title = "Dose-Response Posterior Draws",
# xlab = "Log[Molar]", ylab = "Response"
```






