---
title: "Tidyverse Tutorial"
output: html_notebook
---

If tidyverse and/or tidymodels is not installed yet, you can install the 
packages by using `install.packages("tidyverse")`.

First, we need to load the tidyverse (or tidymodels) library. Uncomment one of 
the libraries and run the code block. 
```{r}
# library(tidyverse)
library(tidymodels)
```


Now, we are going to load in the "bayesetta_tidyverse_dr_data.csv" using 
readr::read_csv()
```{r}
dr_data_orig <- readr::read_csv("C:/Users/marti/Documents/Bayesetta_Workshop/tidyverse_materials/bayesetta_tidyverse_dr_data.csv")
dr_data <- dr_data_orig
dr_data
```

We can see in the data set there's some inconsistencies: 
- there's lower and upper case letters for drug
- the temp_C can be rounded to 0 decimal places
- and the date is not in proper datetime format


Let's start with fixing the drug column so all letters are uppercase. 
- Pipe the data's drug column to `str_replace()`
- Define the pattern and replacement (hint: use toupper as the replacement 
   variable) 
- Assign that to the drug column and display `dr_data` (type dr_data on the 
   last line)

```{r}
dr_data$drug <- dr_data$drug %>%
  stringr::str_replace(pattern = "[:lower:]", replacement = toupper)

dr_data
```

Change temp_C to have 2 significant figures using the `signif()` function. 

```{r}
dr_data$temp_C <- dr_data$temp_C %>%
  signif(2)
dr_data
```

Use `readr::parse_date()` to fix the date format. Use `format = "%m/%d/%Y"`

```{r}
dr_data$date <- dr_data$date %>%
  readr::parse_date("%m/%d/%Y")
dr_data
```

To practice more tidyverse, let's use `mutate()` the data.frame to add a 
`percent_response` column with sigfigs to one decimal place and relocate 
(`relocate()`) it to before "normalized_response".  

```{r}
dr_data <- dr_data %>%
  mutate(percent_response = signif(dr_data$normalized_response*100, 3)) %>%
  relocate(percent_response, .before = normalized_response)
dr_data
```

Let's group by "drug" and use `summarise()` to see the min, max percent 
response, and number of observations (`n()`).
```{r}
dr_data %>% group_by(drug) %>%
  summarise(min_perc_resp = min(percent_response), 
            max_perc_resp = max(percent_response),
            n_observations = n())
```

Let's look at just drug A's dose and percent response in order of decreasing 
dose using `filter()`, `select()`, and `arrange()`.
Assign this to a variable named `dr_drug_A`.

```{r}
dr_drug_A <- dr_data %>% 
  filter(drug == "A") %>%
  select(drug, dose, percent_response) %>%
  arrange(desc(dose))
dr_drug_A
```

Now let's make a simple ggplot (`ggplot()`) of the drug A data points 
(`geom_point()`). You will need to add `scale_x_log10()` for the best display of 
the data points and plot. Try adding a theme (`theme_bw()`), labels (`labs()`), 
and changing colors, point and text size, etc.!
```{r}
ggplot(dr_drug_A) +
  theme_bw() +
  labs(title = "Drug A Dose-Response",
       caption = "this is a simulated dataset",
       y = "Response (%)",
       x = "Dose (M)") +
  theme(axis.text = element_text(size = 12)) +
  geom_point(aes(x = dose, y = percent_response),
             size = 4, color = "magenta") +
  scale_x_log10()
```

Now lets plot all the drugs from the data and use `facet_wrap()` to separate 
each drug into their own plot.
```{r}
ggplot(dr_data) +
  theme_bw() +
  labs(title = "Drug Dose-Response",
       caption = "this is a simulated dataset",
       y = "Response (%)",
       x = "Dose (M)") +
  facet_wrap(vars(drug)) +
  geom_point(aes(x = dose, y = percent_response),
             size = 1, color = "turquoise3") +
  scale_x_log10()
  
```

Save the cleaned data using `readr::write_csv()`. first put the data name then your 
local path for the workshop and a new file name 
(i.e. "clean_bayesetta_dr_data.csv").

```{r}
readr::write_csv(dr_data, 
                 "local_path/Bayesetta_Workshop/dose_response_modeling/clean_bayesetta_dr_data.csv")
```

