---
title: "Tidyverse Tutorial"
output: html_notebook
---

If tidyverse and/or tidymodels is not installed yet, you can install the 
packages by using `install.packages("tidyverse")`.

First, we need to load the tidyverse (or tidymodels) library. Uncomment one of 
the libraries and run the code block. 
```{r}
# library(tidyverse)
# library(tidymodels)
```


Now, we are going to load in the "bayesetta_tidyverse_dr_data.csv" using 
`readr::read_csv()`
```{r}
dr_data_orig <- readr::read_csv("path/to/local/workshop_directory/bayesetta_tidyverse_dr_data.csv")
dr_data <- dr_data_orig
dr_data
```

We can see in the data set there's some inconsistencies: 
- there's lower and upper case letters for drug
- the temp_C can be rounded to 0 decimal places
- and the date is not in proper datetime format


Let's start with fixing the drug column so all letters are uppercase. 
- Pipe the data's drug column to `str_replace()`
- Define the pattern and replacement (hint: use toupper as the replacement 
   variable) 
- Assign that to the drug column and display `dr_data` (type dr_data on the 
   last line)

```{r}
dr_data$drug <- dr_data$drug %>%
  ...

dr_data
```

Change temp_C to have 2 significant figures using the `signif()` function. 

```{r}
 
```

Use `readr::parse_date()` to fix the date format. Use `format = "%m/%d/%Y"`

```{r}
 
```

To practice more tidyverse, let's use `mutate()` the data.frame to add a 
`percent_response` column with sigfigs to one decimal place and relocate 
(`relocate()`) it to before "normalized_response".  

```{r}
 
```

Let's group by "drug" and use `summarise()` to see the min, max percent 
response, and number of observations (`n()`).

```{r}

```

Let's look at just drug A's dose and percent response in order of decreasing 
dose using `filter()`, `select()`, and `arrange()`. 
Assign this to a variable named `dr_drug_A`.

```{r}

```

Now let's make a simple ggplot (`ggplot()`) of the drug A data points 
(`geom_point()`). You will need to add `scale_x_log10()` for the best display of 
the data points and plot. Try adding a theme (`theme_bw()`), labels (`labs()`), 
and changing colors, point and text size, etc.!

```{r}
ggplot(dr_drug_A) +
  geom_point() + 
  ...
```


Now lets plot all the drugs from the data and use `facet_wrap()` to separate 
each drug into their own plot.

```{r}

```

There's some of the basics of tidyverse! We will use this dataset for dose
response modeling later!

Save the cleaned data using `readr::write_csv()`. first put the data name then 
your local path for the workshop and a new file name 
(i.e. "clean_bayesetta_dr_data.csv").

```{r}

```


---

##Plate Reader Example

Testing various concentrations of fluconazole of C. auris. There is one 
replicate (2 columns) of wild type and 2 mutants (ace2 and ace2_ACE2) 
of C.auris. 

Load the excel file using `readxl::read_excel()`. Because of how the columns are 
structured in the excel file, you will probably want to rename the columns of 
the replicate. You can either use `dplyr::rename()` or add a vector of 
`col_names` when loading the excel file. I recommend simply adding an _(number) 
or _rep to the strain name. 

```{r}

```

Now we need to get the tibble (tidyverse data.frame) in a structure better for
analysis. 

Here we will use `tidyr::pivot_longer()`. the `cols` argument will be all the
strains You can either type all the column headers for the strains or exclude
the Fluconazole column by adding an `!` in front of the Fluconazole column 
name. Then, create a column name for `name_to` and `values_to`.

```{r}

```

As we continue the analysis, having different name for the replicate is not 
ideal for plotting and modeling.
To remove the added numbers or letters to the name of the replicate, we can use
dplyr and stringr functions to remove them. 

1. Use mutate to change the strain column or create a new column. 
2. Use `stringr::str_extract()` on the strain column and for the pattern use
a regular expression to extract all the strain names.
3. Pipe the `str_extract()` to `str_remove()` and include the pattern you want
removed from the strain names (i.e. "_1" or _"rep").

```{r}

```

Now we need to convert the Fluconazole values from ug/mL to molar concentration. 
The molecular weight of Fluconazole is 306.271 g/mol. 

Add a new column with the converted values for Fluconazole. 

```{r}
# ug/mL * (10-6g / 1ug) * (1000mL/ 1L) / 306.271 g/mol = M


```


Now let's create a couple plots to see what the data looks like! 

Do one plot with each strain as a different color and one plot using 
`facet_wrap()`

```{r}


```


```{r}

```
